---
title: "Week01 Recitation"
author: "Sophie Kim"
date: "January 28th, 2016"
output: html_document
---


x M- rnorm(100): generates random numbers from the normal distribution

mean(x): mean function

median(x)

var(x)

table(x): how many times each data occurs in a dataset

sort(table(x), decreasing = T): sorts the frequency (used as mode function --> not mode(x))


#Three elements of a function:
1. function
2. argument
3. code block


```{r}

a <- function (zerp) {
  return (zerp)
}




```


prod(x): computes the product of all the elements in vector x

length(x): returns the length of the vector/list

Geometric mean: take nth sqrt of product of n data

--> nice way of getting estimates for skewed data

```{r}

gMean <- function(x) {
  result <- prod(x) ^ 1/(length(x))
  return(result)
}

```

Functions with multiple arguments
```{r}

gMean1 <- function(x, y) {
  return (x + y)
}

```


Functions with default values
```{r}

gMean2 <- function(x, y = 0) {
  return (x + y)
}
```




-----------------------------------------------------------------------------------------------

#Exercise


```{r}

anz <- read.csv("anscombes.csv")

is(anz)

anz[3]

anz[3,]

anz[3:8,]

anz[1:11,]

anz[anz$dataset == 1, ]

anz[anz$dataset == 1, ]$x

anz[anz$dataset == 1, "x"]

anz[anz$dataset == 1, 1]

mean(anz[anz$dataset == 1, 1])


mean(anz[anz$dataset == 1, "x"])

var(anz[anz$dataset == 1, "x"])

mean(anz[anz$dataset == 1, "y"])

var(anz[anz$dataset == 1, "y"])


hist( rnorm(1000))

boxplot( rnorm(1000))

abline(v = 1)

abline(v = 1.5)

abline(v = 2, col = "red")
```

*create a plot first (plot has to be currently active) then use the abline function