---
title: "Week02 Assignment"
author: "Sophie Kim"
date: "February 13, 2016"
output: html_document
---
----------------------------------------------------------------------------------------

#1. Some human phenotypic distributions

1.1 Loading in data
```{r}
bmxData <- read.delim("bmx.tsv")

hist(bmxData$BMXHT, main = "Histogram of Standing Heights", xlab = "Standing Height (cm)", ylab = "Frequency")

```


*the variable BMXHT measures the standing height in cm.


#2. Chi-square test

2.1 Setting up for Mendelian

- Reading the data
```{r}
mendelData <- read.delim("mendel.tab")

```

- Functions to calculate expected values and chi square statistics
```{r}

getExp <- function(proportion, n) {
  expected <- proportion*n
  return (expected)
}

calChi <- function(observed, expected) {
  result <- ((observed - expected)^2)/expected
  return (result)
}



mainCalc <- function(tasknum) {
  n <- mendelData[tasknum, "F2.counts.1st.phenotype"] + mendelData[4, "F2.counts.2nd.phenotype"]
  
  observed1 <- mendelData[tasknum, "F2.counts.1st.phenotype"]
  expected1 <- getExp(0.5, n)
  chi1 <- calChi(observed1, expected1)

  observed2 <- mendelData[tasknum, "F2.counts.2nd.phenotype"]
  expected2 <- getExp(0.5, n)
  chi2 <- calChi(observed2, expected2)
  return (chi1 + chi2)
}

```

1) Round X Wrinkled Seeds
```{r}

chiSum1 <- mainCalc(1)
chiSum1

pchisq(chiSum1, 1)
```


2) Yellow X Green Seeds
```{r}

chiSum2 <- mainCalc(2)
chiSum2

pchisq(chiSum2, 1)
```


3) Purple X White Seeds
```{r}

chiSum3 <- mainCalc(3)
chiSum3

pchisq(chiSum3, 1)
```


4) Inflated X Pinched Pods
```{r}


chiSum4 <- mainCalc(4)
chiSum4

pchisq(chiSum4, 1)
```


5) Green X Yellow Pods
```{r}

chiSum5 <- mainCalc(5)
chiSum5

pchisq(chiSum5, 1)
```


6) Axial X Terminal Flowers
```{r}

chiSum6 <- mainCalc(6)
chiSum6

pchisq(chiSum6, 1)
```


7) Long X Short Stems
```{r}
chiSum7 <- mainCalc(7)
chiSum7

pchisq(chiSum7, 1)

```


2.2 3:1 ratios


- Functions to calculate expected values and chi square statistics
```{r}

getExp <- function(proportion, n) {
  expected <- proportion*n
  return (expected)
}

calChi <- function(observed, expected) {
  result <- ((observed - expected)^2)/expected
  return (result)
}



mainCalc <- function(tasknum) {
  n <- mendelData[tasknum, "F2.counts.1st.phenotype"] + mendelData[4, "F2.counts.2nd.phenotype"]
  
  observed1 <- mendelData[tasknum, "F2.counts.1st.phenotype"]
  expected1 <- getExp(0.75, n)
  chi1 <- calChi(observed1, expected1)

  observed2 <- mendelData[tasknum, "F2.counts.2nd.phenotype"]
  expected2 <- getExp(0.25, n)
  chi2 <- calChi(observed2, expected2)
  return (chi1 + chi2)
}

```

1) Round X Wrinkled Seeds
```{r}

chiSum1 <- mainCalc(1)
chiSum1

pchisq(chiSum1, 1)

```


2) Yellow X Green Seeds
```{r}

chiSum2 <- mainCalc(2)
chiSum2

pchisq(chiSum2, 1)
```


3) Purple X White Seeds
```{r}

chiSum3 <- mainCalc(3)
chiSum3

pchisq(chiSum3, 1)
```


4) Inflated X Pinched Pods
```{r}


chiSum4 <- mainCalc(4)
chiSum4

pchisq(chiSum4, 1)
```


5) Green X Yellow Pods
```{r}

chiSum5 <- mainCalc(5)
chiSum5

pchisq(chiSum5, 1)
```


6) Axial X Terminal Flowers
```{r}

chiSum6 <- mainCalc(6)
chiSum6

pchisq(chiSum6, 1)
```


7) Long X Short Stems
```{r}
chiSum7 <- mainCalc(7)
chiSum7

pchisq(chiSum7, 1)
```


2.3 Some more Mendelian Genetics

Imagine a cross between two diploid individuals heterozygous for ten independently assorting Mendelian traits. 

1) What is the probability of one of their progeny being homozygous recessive for all traits? 
```{r}
(1/16)^10
```

2)What is the probability of a progeny that is homozygous dominant for all traits?
```{r}
(1/16)^10
```



#3. and Bateson et al observe coupling

1. Reading Bateson tab data
```{r}
bData <- read.delim("bateson.tab", header = FALSE)

```
- factors: shape, color

- levels: purple, red, long, round

- degrees of freedom = 1

```{r}
calChi <- function(observed, expected) {
  result <- ((observed - expected)^2)/expected
  return (result)
}

n <- sum(bData[,2])

purple <- bData[1,2] + bData[3,2]
long <- bData[1,2] + bData[1,2]
red <- bData[2,2] + bData[4,2]
round <- bData[2,2] + bData[4,2]

PL <- (purple*long)/n
pL <- (red*long)/n
Pr <- (purple*round)/n
pr <- (red*round)/n

chisquareSum <- calChi(bData[1,2], PL) + calChi(bData[2,2], pL) + calChi(bData[3,2], Pr) + calChi(bData[4,2], pr)

chisquareSum

```

*there was a minor type-o in accessing the bData[1,2]